```{r}
usgs_gages <- read.csv(here("1-data/instrumented_all-predictors.csv"))
model_data <- read.csv(here("2-experiments/exp03_trimmed-statewide-model/data/trimmed-statewide-model_results.csv"))
```

## Produce table with instrumented gauges used in study

To be added to Supplemental Information

```{r}
summary_df <- usgs_gages %>%
  group_by(Site_Num, HUC8, Lat, Long) %>%
  summarise(
    num_years_of_record = n(),
    years_of_record = list(sort(unique(Year)))
  ) %>%
  ungroup()

summary_df$years_of_record <- sapply(summary_df$years_of_record, toString)
# write.csv(summary_df, here("1-data/instrumented_period-of-record.csv"), row.names = F)
```

## Loading Initial USGS Streamgages

```{r}
# Get discharge data for all AZ streamgages
sites_init <- dataRetrieval::whatNWISsites(stateCd = "AZ", #Only sites in AZ
                            parameterCd = "00060",  #discharge
                            service="dv")

# Preprocessing retrieved streamgages
sites <- filter(sites_init, site_tp_cd %in% c('ST')) #only sites that are a Stream
sites <- sites[,-c(1,4,7,8)] #remove unwanted columns
colnames(sites) = c('Site_Num', 'Station_Name', 'Latitude', 'Longitude')
sites <- sites[c(1:254,256:478),] #hard code remove specific problem site

#################################################
## Remove Colorado River gauges
#################################################

# Find indices where "Colorado River" appears in the Station_Name column
colorado_sites_indices <- grep("Colorado River", sites$Station_Name, ignore.case = TRUE)

little_colorado_sites_indices <- grep("Little Colorado River", sites$Station_Name, ignore.case = TRUE)

# Exclude "Little Colorado River" from the indices of "Colorado River"
final_colorado_sites_indices <- setdiff(colorado_sites_indices, little_colorado_sites_indices)

sites_noCO <- sites[-final_colorado_sites_indices, ]
```

# Filtering Streamgages 

### Initialize dataframe

This dataframe will hold the goodness of fit statistics for the model on each classification of streamgages

```{r}
stats.Class <- data.frame()
```

## Seasonality

Monsoon-dominated: Majority of streamflow occurs during July-September

Snowmelt-dominated: Majority of streamflow occurs during March-June

```{r}
source(here("1-data/preprocessing/1_preprocess_fx.R"))

USGSsites = annualUSGS_preprocessing(sites_noCO)
yearly_bfi <- USGSsites$results

seasonality_count <- yearly_bfi %>%
  group_by(Site_Num, Seasonality) %>%
  summarise(count = n()) %>%
  ungroup()
  
seasonality_gauges <- seasonality_count %>%
  group_by(Site_Num) %>%
  slice_max(count, n = 1, with_ties = FALSE) %>%
  dplyr::select(Site_Num, Seasonality)
```

### Model Performance

```{r}
seasonality_gauges$Site_Num <- as.integer(seasonality_gauges$Site_Num) #both are integers

# Left join to add 'Seasonality' column to the dataframe
data.seasonal <- model_data %>%
  left_join(seasonality_gauges, by = "Site_Num")

season.Monsoon <- which(data.seasonal$Seasonality == "Monsoon-dominated")
season.Snow <- which(data.seasonal$Seasonality == "Snowmelt-dominated")

# Monsoon-dominated catchment Goodness of Fit Stats
monsoon.stat <- c(Class = "Climate - Monsoon Dominated", analysis.stats(data.seasonal[season.Monsoon,]))
stats.Class <- rbind(stats.Class, monsoon.stat)

# Snowmelt-dominated catchment Goodness of Fit Stats
snow.stat <- c(Class = "Climate - Snowmelt Dominated", analysis.stats(data.seasonal[season.Snow,]))
stats.Class <- rbind(stats.Class, snow.stat)

```

## Physiographic Region

```{r}
regional_inst.data <- read.csv(here("1-data/raw/instrumented_region.csv"))

streamgage_regional.data <- regional_inst.data %>%
  distinct(Site_Num, .keep_all = TRUE) #unique usgs streamgages only
streamgage_regional.data <- streamgage_regional.data[,c(1:4,6)]
```

### Model Performance

```{r}
data.regional <- model_data %>%
  left_join(streamgage_regional.data, by = "Site_Num") #add PROVINCE column to dataset

region.Basin <- which(data.regional$PROVINCE == "BASIN AND RANGE")
region.Plateau <- which(data.regional$PROVINCE == "COLORADO PLATEAUS")

# Basin&Range Goodness of Fit Stats
basin.stat <- c(Class = "PhysRegion - Basin&Range", analysis.stats(data.seasonal[region.Basin,]))
stats.Class <- rbind(stats.Class, basin.stat)

# Colorado Plateau Goodness of Fit Stats
plateau.stat <- c(Class = "PhysRegion - CO Plateau", analysis.stats(data.seasonal[region.Plateau,]))
stats.Class <- rbind(stats.Class, plateau.stat)
```

## Climate

Warm-Wet: above median temp, above median precip

Warm-Dry: above median temp, below median precip

Cool-Wet: below median temp, above median precip

Cool-Dry: below median temp, below median precip

```{r}
data.climate <- model_data %>%
  mutate(Precip_Category = ifelse(Precip_MM <= median(Precip_MM), "Dry", "Wet"))

# Create a new column to categorize temperature based on median
data.climate <- data.climate %>%
  mutate(Temp_Category = ifelse(Temp_C <= median(Temp_C), "Cool", "Warm"))

# Create the final climate groups using the "Warm/Cool-Wet/Dry" terminology
data.climate <- data.climate %>%
  mutate(Climate_Group = case_when(
    Precip_Category == "Dry" & Temp_Category == "Cool" ~ "Cool-Dry",
    Precip_Category == "Dry" & Temp_Category == "Warm" ~ "Warm-Dry",
    Precip_Category == "Wet" & Temp_Category == "Cool" ~ "Cool-Wet",
    Precip_Category == "Wet" & Temp_Category == "Warm" ~ "Warm-Wet"
  ))

```

### Model Performance

```{climate_WW <- which(data.climate$Climate_Group == "Warm-Wet")}
climate_WD <- which(data.climate$Climate_Group == "Warm-Dry")
climate_CW <- which(data.climate$Climate_Group == "Cool-Wet")
climate_CD <- which(data.climate$Climate_Group == "Cool-Dry")

# Warm-Wet Goodness of Fit Stats
climate_WW.stat <- c(Class = "Climate - Warm-Wet", analysis.stats(data.climate[climate_WW,]))
stats.Class <- rbind(stats.Class, climate_WW.stat)

# Warm-Dry Goodness of Fit Stats
climate_WD.stat <- c(Class = "Climate - Warm-Dry", analysis.stats(data.climate[climate_WD,]))
stats.Class <- rbind(stats.Class, climate_WD.stat)

# Cool-Wet Goodness of Fit Stats
climate_CW.stat <- c(Class = "Climate - Cool-Wet", analysis.stats(data.climate[climate_CW,]))
stats.Class <- rbind(stats.Class, climate_CW.stat)

# Cool-Dry Goodness of Fit Stats
climate_CD.stat <- c(Class = "Climate - Cool-Dry", analysis.stats(data.climate[climate_CD,]))
stats.Class <- rbind(stats.Class, climate_CD.stat)
```

## Slope Gradient

```{r}
# Calculate the slope gradient for each site and add as a new column
data.slope <- usgs_gages %>%
  mutate(Slope_Decimal = tan(SLOPE_DEGREES * pi / 180))  # Convert slope from degrees to decimals (slope gradient)

# Categorize each site into "Low" or "High" based on the median slope
data.slope <- data.slope %>%
  mutate(Slope_Category = ifelse(Slope_Decimal <= median(Slope_Decimal), "Low", "High"))

temp_data.slope <- data.slope %>% 
  dplyr::select(Site_Num, Slope_Category) %>%
  distinct(Site_Num, Slope_Category)

# Add slope to model_data
slope_data <- left_join(model_data, temp_data.slope, by = "Site_Num")
```

```{r}
slope_high <- which(slope_data$Slope_Category == "High")
slope_low <- which(slope_data$Slope_Category == "Low")

# Steep Slope Goodness of Fit Stats
slope_high_stat <- c(Class = "Slope - High", analysis.stats(slope_data[slope_high, ]))
stats.Class <- rbind(stats.Class, slope_high_stat)

# Gentle Slope Goodness of Fit Stats
slope_low_stat <- c(Class = "Slope - Low", analysis.stats(slope_data[slope_low, ]))
stats.Class <- rbind(stats.Class, slope_low_stat)
```
