```{r}
usgs_gages <- read.csv(here("1-data/instrumented_all-predictors.csv"))
```

## Step 1: Descriptive Statistics for BFI

Calculate basic descriptive statistics for BFI: mean, median, standard deviation, min, max, and quartiles

-   \>400 observations of BFI are \< .0001; 1000 are \< 0.01

    -   `Should check if they are at the same locations (rivers to get rid of as they are likely ephemeral) or time periods (dry periods of groundwater-surface water interaction)`

```{r}
bfi_summary <- usgs_gages %>%
  summarise(
    mean_BFI = mean(BFI, na.rm = TRUE),    # Mean BFI
    median_BFI = median(BFI, na.rm = TRUE),  # Median BFI
    sd_BFI = sd(BFI, na.rm = TRUE),        # Standard Deviation of BFI
    min_BFI = min(BFI, na.rm = TRUE),      # Minimum BFI
    max_BFI = max(BFI, na.rm = TRUE),      # Maximum BFI
    Q1_BFI = quantile(BFI, 0.25, na.rm = TRUE),  # First Quartile (25th percentile)
    Q3_BFI = quantile(BFI, 0.75, na.rm = TRUE)   # Third Quartile (75th percentile)
  )

bfi_summary

# Plots of BFI
ggplot(usgs_gages, aes(x = BFI)) + 
  geom_histogram(binwidth = 0.01, color = 'black', fill = 'purple', alpha = 0.7) + 
  labs(title = "Histogram of BFI", x = "Baseflow Index (BFI)", y = "Frequency") + 
  theme_minimal()

length(which(usgs_gages$BFI < .01))
length(which(usgs_gages$BFI < .001))

```

## Step 2: Kendall Tau Analysis

Kendall's Tau correlation test for BFI with other numeric variables

-   **No strong correlations** were found between BFI and any of the variables tested

-   The relationship between **Year** and **BFI** is negligible (0.025, weak positive correlation), indicating no trend over time for BFI in the data.

```{r}
library(corrr)

# Select the desired columns from the dataset
select_columns <- usgs_gages %>%
  dplyr::select(BFI, Year, Temp_C, Precip_MM, AET_MM, Elevation_M, AREA_KM2)

# Compute Kendall's Tau correlations
kendall_tau_corr <- cor(select_columns, use = "pairwise.complete.obs", method = "kendall")

# Extract BFI correlations
bfi_corr <- kendall_tau_corr["BFI", ]

# Correlation matrix as a heatmap
library(ggcorrplot)
ggcorrplot(kendall_tau_corr, lab = TRUE, title = "Kendall's Tau Correlation Matrix")

```

## Multiple Linear Regression Model

```{r}
# Fit a multiple linear regression model
model <- lm(BFI ~ ., data = usgs_gages)

# summary of the model
summary(model)

# use model to make predictions
predictions <- predict(model, usgs_gages)

# calculate model performance metrics
r2_value <- summary(model)$r.squared
mse_value <- mse(usgs_gages$BFI, predictions)
rmse_value <- rmse(usgs_gages$BFI, predictions)
mae_value <- mae(usgs_gages$BFI, predictions)
nse_value <- NSE(obs = usgs_gages$BFI, sim = predictions)
pbias_value <- (sum(usgs_gages$BFI - predictions) / sum(usgs_gages$BFI)) * 100

results <- data.frame(
    "R2" = r2_value,
    "MSE" = mse_value,
    "RMSE" = rmse_value,
    "MAE" = mae_value,
    "Nash-Sutcliffe" = nse_value,
    "pbias" = pbias_value)

results

```

## Exploring Low BFI - Spatially/Temporally

```{r}
# Filter observations of BFI < 0.0001
low_bfi_data <- usgs_gages %>% filter(BFI < 0.0001)

# Count occurrences of low BFI observations by location
location_counts <- low_bfi_data %>%
  group_by(Site_Num, HUC8) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# View locations with low BFI observations
print(location_counts)

```
